---
title: "Figure 3"
---

```{r knitr, include = FALSE}
DOCNAME = "figure-3"
NOW <- Sys.time()

# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
    if (before) {
        print(paste("Start:", Sys.time()))
        NOW <<- Sys.time()
    } else {
        print(paste("Stop:", Sys.time()))
        print(Sys.time() - NOW)
    }
})

knitr::opts_chunk$set(
    autodep        = TRUE,
    cache          = TRUE,
    cache.lazy     = FALSE,
    cache.path     = paste0("cache/", DOCNAME, "/"),
    cache.comments = FALSE,
    echo           = TRUE,
    error          = FALSE,
    fig.align      = "center",
    fig.width      = 10,
    fig.height     = 8,
    message        = FALSE,
    warning        = FALSE,
    timeit         = TRUE
)
```

```{r libaries, cache = FALSE}
# scRNA-seq
library(Seurat)
#packageVersion("Seurat")

# Plotting
library("ggplot2")

# Presentation
library("knitr")

# Others
library("rstatix")
library('scales')
library("readxl")
```

```{r source, cache = FALSE}
source(here::here("R/00_generalDeps.R"))
source(here::here("R/output.R"))
```

<!-- # ```{r depends-paths} -->
<!-- # filt_path <- here::here("data/processed/AmR053_harmony_azimuth_V4regRB.rds") -->
<!-- # ``` -->

Introduction
============

In this document we are going to perform clustering and plotting figure 3 panel B C & E,  Supplementary figure 5 and Supplementary figure 6
using `Seurat`.


Seurat -Integration by harmony
==============

Parse TCR
------

```{r parse_TCR}
# 	# Function parser TCR cellRanger output
#   # Add Clonotypes functions
#    add_clonotype <- function(tcr_prefix, seurat_obj, type,num){
#     vdj_file=file.path(tcr_prefix,"filtered_contig_annotations.csv")
#     if(file.exists(vdj_file)){ 
#       res <- tryCatch({
#         tcr <- read.csv(vdj_file)
#         
#         # Extract all informations of V(d)J and c genes per barcode
#         if(type =='b'){
#           tcr <- as.data.frame(format_b_line(tcr))
#         } else if (type == 't'){
#           tcr <- as.data.frame(format_t_line(tcr))
#         }
#         
#         # Clonotype-centric info.
#         clono <- read.csv(paste(tcr_prefix,"clonotypes.csv", sep="/"))
#         clono <- clono[, c("clonotype_id", "cdr3s_aa", "proportion")]
#         colnames(clono) <- paste(type, colnames(clono), sep="_")
#         # Slap the AA sequences onto our original table by clonotype_id.
#         tcr <- merge(tcr, clono,by.x='clonotype_ID', by.y=paste(type,'clonotype_id',sep='_'))
#         row.names(tcr) <- paste(tcr$barcode,'_',num,sep='')
#         tcr$barcode <- NULL
#         # Add to the Seurat object's metadata.
#         clono_seurat <- AddMetaData(object=seurat_obj, metadata=tcr)
#         return(clono_seurat)
#       }, error =  function(e) {
#         print(e)
#         return(seurat_obj)
#       })
#     } else {
#       print(paste('Cannot find file : ',vdj_file))
#     }
#     return(seurat_obj)
#   }
#   
#   #Function format TCR output
#   format_t_line <- function(tcr){
#     list_barcode <- unique(tcr$barcode)
#     vdj_list <- c()
#     for(barcode in unique(tcr$barcode)){
#       list_bc <- tcr[tcr$barcode==barcode,]
#       clonoID <- unique(list_bc$raw_clonotype_id)
#       TCRID <- paste(paste(list_bc$chain,paste(list_bc$v_gene,list_bc$d_gene,list_bc$j_gene,list_bc$c_gene,sep=';'),sep=':'),collapse=' ')
#       TR_list <- list()
#       TR_list[['TRA']] <- list_bc$chain[list_bc$chain=='TRA']
#       TR_list[['TRAV']] <- list_bc$v_gene[list_bc$chain=='TRA'  & list_bc$v_gene!='' ]
#       TR_list[['TRAD']] <- list_bc$d_gene[list_bc$chain=='TRA' & list_bc$d_gene!='' ]
#       TR_list[['TRAJ']] <- list_bc$j_gene[list_bc$chain=='TRA' & list_bc$j_gene!='' ]
#       TR_list[['TRAC']]<- list_bc$c_gene[list_bc$chain=='TRA' & list_bc$c_gene!='' ]
#       
#       TR_list[['TRB']] <- list_bc$chain[list_bc$chain=='TRB']
#       TR_list[['TRBV']] <- list_bc$v_gene[list_bc$chain=='TRB'  & list_bc$v_gene!='' ]
#       TR_list[['TRBD']] <- list_bc$d_gene[list_bc$chain=='TRB' & list_bc$d_gene!='' ]
#       TR_list[['TRBJ']] <- list_bc$j_gene[list_bc$chain=='TRB' & list_bc$j_gene!='' ]
#       TR_list[['TRBC']]<- list_bc$c_gene[list_bc$chain=='TRB' & list_bc$c_gene!='' ]
#       
#       TR_list_out <- c()
#       names_out <- c()
#       for(tr in names(TR_list)){
#         if(length(TR_list[[tr]])==1){
#           TR_list_out <- c(TR_list_out,TR_list[[tr]][1],NA)
#         } else if (length(TR_list[[tr]])==0){
#           TR_list_out <- c(TR_list_out,NA,NA)
#         } else {
#           TR_list_out <- c(TR_list_out,TR_list[[tr]])
#         }
#         names_out <- c(names_out,paste(tr,c('1','2'),sep='_'))
#         names(TR_list_out) <- names_out
#       }
#       
#       vdj_list <- rbind(vdj_list,c('barcode'=barcode,'TCR_ID'=TCRID,TR_list_out,'clonotype_ID'=clonoID))
#       
#     }
#     return(vdj_list)
#   }
```

Load matrices
------

```{r load_matrices}
# # Load Matrices and remove TCR genes from it
# mat_v1<-Read10X(data.dir="CellRanger/AmR053_V1_Multi/sample_filtered_feature_bc_matrix",gene.column=2,unique.features=TRUE)
# colnames(mat_v1)<-paste(colnames(mat_v1),'1',sep='_')
# mat_v1 <- mat_v1[grep("^TRA[JV]|^TRB[JV]|^TRG[JV]|^TRD[JV]",row.names(mat_v1),invert=TRUE),]
# mat_v2<-Read10X(data.dir='CellRanger/AmR053_V2_Multi/sample_filtered_feature_bc_matrix',gene.column=2,unique.features=TRUE)
# colnames(mat_v2)<-paste(colnames(mat_v2),'2',sep='_')
# mat_v2 <- mat_v2[grep("^TRA[JV]|^TRB[JV]|^TRG[JV]|^TRD[JV]",row.names(mat_v2),invert=TRUE),]
# mat_v3<-Read10X(data.dir='CellRanger/AmR053_V3_Multi/sample_filtered_feature_bc_matrix',gene.column=2,unique.features=TRUE)
# colnames(mat_v3)<-paste(colnames(mat_v3),'3',sep='_')
# mat_v3 <- mat_v3[grep("^TRA[JV]|^TRB[JV]|^TRG[JV]|^TRD[JV]",row.names(mat_v3),invert=TRUE),]
# mat_v4<-Read10X(data.dir='CellRanger/AmR053_V4_Multi/sample_filtered_feature_bc_matrix',gene.column=2,unique.features=TRUE)
# colnames(mat_v4)<-paste(colnames(mat_v4),'4',sep='_')
# mat_v4 <- mat_v4[grep("^TRA[JV]|^TRB[JV]|^TRG[JV]|^TRD[JV]",row.names(mat_v4),invert=TRUE),]
```


Setup the Seurat Object
------

```{r Seurat_obj}
# # Create Seurat objects from matrices and add TCR metadata
# amr_v1 <- CreateSeuratObject(counts = mat_v1, project = "AmR053_V1", min.cell=3,min.features=200)
# amr_v1$samples <- 'V1'
# amr_v1 <- add_clonotype('CellRanger/AmR053_V1_Multi/vdj_t', amr_v1, "t",'1')
# amr_v2 <- CreateSeuratObject(counts = mat_v2, project = "AmR053_V2", min.cell=3,min.features=200)
# amr_v2$samples <- 'V2'
# amr_v2 <- add_clonotype('CellRanger/AmR053_V2_Multi/vdj_t', amr_v2, "t",'2')
# amr_v3 <- CreateSeuratObject(counts = mat_v3, project = "AmR053_V3", min.cell=3,min.features=200)
# amr_v3$samples <- 'V3'
# amr_v3 <- add_clonotype('CellRanger/AmR053_V3_Multi/vdj_t', amr_v3, "t",'3')
# amr_v4 <- CreateSeuratObject(counts = mat_v4, project = "AmR053_V4", min.cell=3,min.features=200)
# amr_v4$samples <- 'V4'
# amr_v4<- add_clonotype('CellRanger/AmR053_V4_Multi/vdj_t', amr_v4, "t",'4')
# amr_list <- list('v1'=amr_v1, 'v2'=amr_v2,'v3'=amr_v3,'v4'=amr_v4)
```

Integration
------

```{r integration}
# for(v in names(amr_list)){
#         # Remove doublet based on TCR annotation
#         #amr_list[[v]][['doubletTCR']] <- !(is.na(amr_list[[v]][['TRA_2']]) & is.na(amr_list[[v]][['TRB_2']]))
#         # Add mitochondrial genes percentage
#         amr_list[[v]][['percent.mt']] <- PercentageFeatureSet(amr_list[[v]],pattern="^[MTmt]{2}-")
#         # Add rebosomal genes percentage
#         amr_list[[v]][['percent.rb']] <- PercentageFeatureSet(amr_list[[v]],pattern="^RP[SL][[:digit:]]|^RPLP[[:digit:]]|^RPSA")
#         # Subset object on number of genes per cells and TCR doublet
#         amr_list[[v]] <-subset(amr_list[[v]],subset=nFeature_RNA > 500 & nFeature_RNA < 4000 & percent.mt < 5)# & doubletTCR==FALSE)
#         # Normalise counts and regress on precentage of ribosomal genes
#         amr_list[[v]] <- SCTransform(amr_list[[v]], vars.to.regress = "percent.rb", verbose = FALSE)
# }
# # Calculate integration features
# integ_feat <- SelectIntegrationFeatures(amr_list, nfeatures=3000)
# # Merge the four matrices
# amr_all <- merge(amr_list[['v1']], y=c(amr_list[['v2']],amr_list[['v3']],amr_list[['v4']]), project ='AMR053')
# # Integrate all data set with harmony
# VariableFeatures(amr_all) <- integ_feat
# amr_all <- RunPCA(amr_all, assay = "SCT", npcs = 50)
# amr.harmony <- RunHarmony(amr_all,"samples", assay.use='SCT',plot_convergence = FALSE)
```

Clustering
------

```{r UMAP}
# # Calculate clusters with resolution of 0.3
# reso=0.3
# amr.harmony <- FindNeighbors(amr.harmony, reduction = "harmony", dims = 1:13)
# amr.harmony <- FindClusters(amr.harmony, resolution = reso,graph.name='SCT_snn')
# amr.harmony <- RunUMAP(amr.harmony, reduction = "harmony", assay = 'SCT',dims = 1:13,seed.use=42)
# amr.harmony.umap <- DimPlot(amr.harmony, reduction = "umap", group.by = "seurat_clusters", pt.size = .1,ncol=2, label=TRUE, seed=42)
# umaps <- amr.harmony.umap
```

Markers HM
------

```{r markers_HM}
# #Indentify marker genes for each cluster
# DefaultAssay(amr.harmony) <- 'SCT'
# amr.harmony<-ScaleData(amr.harmony, features = row.names(amr.harmony))
# amr.harmony <- PrepSCTFindMarkers(amr.harmony, assay = "SCT", verbose = TRUE)
# amr.harmony.mkrs <-  FindAllMarkers(amr.harmony, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
# amr.harmony.mkrs_padj <- amr.harmony.mkrs[amr.harmony.mkrs$p_val_adj <0.01 ,]
# amr.harmony.mkrs_padj %>% group_by(cluster) %>% top_n(n = 5, wt = avg_log2FC) -> top5
# heatmap.mkrs <- DoHeatmap(amr.harmony, features=top5$gene, group.by='seurat_clusters')
```

Markers DP
------

```{r markers_DP}
# list_feat=c('MAL','AQP3','TCF7','CCL5','GZMK','GZMA','HLA-DRA','LGALS1','CXCR3','NKG7','GZMH','PLEK','HIST1H1E','HIST1H1D','COTL1','GNLY','GZMB','PRF1','FOXP3','IKZF2','RTKN2','XIST','NEAT1','MACF1','AL627171.2','PLCG2','MT-ND6')
# dot.mkrs <- DotPlot(amr.harmony,features=list_feat, cols='RdBu', dot.scale=12) + RotatedAxis() + ggtitle("Marker genes") + ylab('Clusters')
```


Module score
------

```{r module_score, echo=FALSE, message=FALSE, warning=FALSE}
# #Genes module autoreactivity
# autoreact <- str_split("MAGEH1,NMB,TIGIT,SPOCK2,GK,CTLA4,ITGA4,NR3C1,SYNE2,CD109,ICA1,ITM2A,TOX,HNRNPLL,DYNLL1,NDUFC1,IGFL2,LAPTM5,TSHZ2,LINC01480,TBC1D4,NIN,ARHGAP26,IFNAR2,CRYBG1,DCXR,CXCL13,RSRP1,TRAT1,TMEM59,TMSB4XP8,HLA-DRB1,AKAP13,NDUFV2P1,HLA-DPB1,STAT3,HIST1H4C,SGK1,MAPRE2,WBP2,RHBDD2,HMGN2,CD59,ANKRD12,MAT2B,PAIP2,ARPC5,FKBP5,HLA-DRB5,RPS27L,ARPC1B,PSMB9,PTTG1,HMGB2,WSB1,LAPTM4A,FAM177A1,ANXA5,SARAF,MIR4435-2HG,IFITM1,LYST,ATP6V1D,YPEL5,ARID5B,ARHGDIA,FABP5,IK,HLA-DPA1,SERINC3,APLP2,CHD2,ITM2B,ATXN1,RGS2,SEC11A,CD74,HMGN3,MPHOSPH8,TMEM50B,S100A4,PSAP,HBP1,DBI,H2AFZ,PPP2R5C,LEPROTL1,BTLA,TAX1BP1,KLRB1,LAMP2,CNIH4,RNASET2,CIRBP,ISG15,CHN1,ANAPC16,SLC20A1,CYTH1,MEA1",',')[[1]]
# amr.harmony <- AddModuleScore(amr.harmony, features=list(autoreact), name='auto_reactivity', search=TRUE)
# #Genes module H1N1
# h1n1 <- str_split("FASLG,GZMB,CCL5,IL3,TOMM5,CSF2,TNFRSF9,CCL4,NME1,RP11-343H5.4,APOBEC3G,RP4-765C7.2,TNFSF14,CAMTA2,RP11-543P15.1,RN7SL1,RGS16,RP4-604A21.1,CARS2,TNFSF10,APOBEC3C,GCNA,FOSL1,NINJ1,IL32,GPR183,CD48,EIF2S2,XCL1,MYC,DDX21,EIF5A,EXOSC8,ACSL6,IL10,KLF10,NT5E,PMAIP1,SNHG8,SNRPGP10,IL23A,RUNX3,BMI1,CCL4L2,HSPE1,BACH2,RP11-490H24.5,SLIRP,PLEK,CCND2,PRMT1,BTG1,YBX1,NKG7,GPATCH4,NOP16,CXCR3,LINC00892,DDX18,IFNG,LYAR,PA2G4,NOLC1,PPA1,XBP1,CBWD1,UTP4,PTPRE,NUCB2,SNHG16,SYNCRIP,AC006369.2,WDR43,ANP32B,CST7,BRIX1,PHLDA1,TNFSF9,RPL26,CBFB,VTRNA1-3,TRAV8-4,EIF5B,SLC43A3,EPRS,NDFIP2,EBNA1BP2,RSL24D1,SEC61G,SRPRB,IRF8,RPL39,ABRACL,MTHFD2,DKC1,GTPBP4,FKBP11,HSP90B1,PSME2,DNTTIP2",',')[[1]]
# amr.harmony <- AddModuleScore(amr.harmony, features=list(h1n1), name='H1N1', search=TRUE)
# 
# # Feature plot auto reactivity highlight top 5%
# top95p_auto <- tail(quantile(t(amr.harmony[[paste('auto_reactivity','1',sep='')]]),seq(0,1,0.05)),2)[1]
# top_cells <- colnames(amr.harmony)[t(amr.harmony[[paste('auto_reactivity','1',sep='')]]) > top95p_auto]
# pt_size_set <- rep(0.01, dim(amr.harmony)[2])
# names(pt_size_set) <- colnames(amr.harmony)
# pt_size_set[top_cells] <- 1
# ft_modlist_auto <-FeaturePlot(amr.harmony, features=paste('auto_reactivity','1',sep=''), label=TRUE, pt.size=pt_size_set) + scale_colour_gradientn(colours=rev(brewer.pal(n=11, name="RdBu"))) + ggtitle("Auto-reactivity module score Top50 highlight top 5%")
# 
# # Feature plot H1N1 highlight top 5%
# top95p_h1n1 <- tail(quantile(t(amr.harmony[[paste('H1N1','1',sep='')]]),seq(0,1,0.05)),2)[1]
# top_cells <- colnames(amr.harmony)[t(amr.harmony[[paste('H1N1','1',sep='')]]) > top95p_h1n1]
# pt_size_set <- rep(0.01, dim(amr.harmony)[2])
# names(pt_size_set) <- colnames(amr.harmony)
# pt_size_set[top_cells] <- 1
# ft_modlist_h1n1 <-FeaturePlot(amr.harmony, features=paste('H1N1','1',sep=''), label=TRUE, pt.size=pt_size_set) + scale_colour_gradientn(colours=rev(brewer.pal(n=11, name="RdBu"))) + ggtitle("H1N1 module score Top50 highlight top 5%")
```

MS violin
------

```{r MS_violin, echo=FALSE, message=FALSE, warning=FALSE}
# # List module score  violin Plot pairwise wilcoxon test
# #AutoReactivity
# liste_auto <- amr.harmony[[c('auto_reactivity1','seurat_clusters')]]
# names(liste_auto) <- c('mod_score','clusters')
# pwWlcx_auto <- pairwise_wilcox_test(liste_auto,formula=mod_score ~ clusters,p.adjust.method = "BH")
# mod_plot_auto <-  ggviolin(liste_auto, x='clusters', y='mod_score',palette=hue_pal()(10), fill='clusters') + 
# stat_pvalue_manual(pwWlcx_auto, hide.ns = FALSE, y.position=0.5, step.increase=0.05) + ggtitle('Auto Reactivity')
# 
# #H1N1
# liste_h1n1 <- amr.harmony[[c('H1N11','seurat_clusters')]]
# names(liste_h1n1) <- c('mod_score','clusters')
# pwWlcx_h1n1 <- pairwise_wilcox_test(liste_h1n1,formula=mod_score ~ clusters,p.adjust.method = "BH")
# mod_plot_h1n1 <-  ggviolin(liste_h1n1, x='clusters', y='mod_score',palette=hue_pal()(10), fill='clusters') + 
# stat_pvalue_manual(pwWlcx_h1n1, hide.ns = FALSE, y.position=0.5, step.increase=0.05) + ggtitle('H1N1')
```

Azimuth
------

```{r azimuth}
# #Cluster annotation with azimuth, load azimuth annotation from previous run v0.4.6
# DefaultAssay(amr.harmony) <- 'SCT'
# Idents(amr.harmony) <- 'seurat_clusters'
# amr.harmony <- RunAzimuth(amr.harmony, reference = "pbmcref", assay='SCT')
# azm_plot <- DimPlot(amr.harmony, reduction='umap', split.by='predicted.celltype.l2',ncol=5,seed=42)
```

Saving
------

```{r saving}
# saveRDS(amr.harmony,file=here::here("data/processed/AmR053_harmony_azimuth_V4regRB.rds"))
```

```{r depends-paths}
filt_path <- here::here("data/processed/AmR053_harmony_azimuth_V4regRB.rds")
```


Loading `seurat` object processed
```{r load, cache.extra = tools::md5sum(filt_path)}

if (file.exists(filt_path)) {
    amr.harmony <- readRDS(filt_path)
} else {
    stop("Figure 3  dataset is missing. ",
         "Please check Input first.",
         call. = FALSE)
}
```


Figure 3
=======

Panel B
------

```{r fig3-pB, fig.width = 7, fig.height = 6}

fig <- DimPlot(amr.harmony, reduction = "umap", seed=42, group.by = "seurat_clusters", pt.size = .1,ncol=1, label=TRUE) + ggtitle(paste0("seurat_clusters: ", length(amr.harmony$seurat_clusters)))

dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

ggsave(here::here("output", DOCNAME, "figure3-panelB.pdf"), fig,
        width = 7, height = 6, scale = 1)
ggsave(here::here("output", DOCNAME, "figure3-panelB.png"), fig,
        width = 7, height = 6, scale = 1)

fig
```

Panel C
------

```{r fig3-pC, fig.width = 9, fig.height = 6}
DefaultAssay(amr.harmony) <- 'SCT'
amr.harmony.mkrs <-  FindAllMarkers(amr.harmony,assay='SCT', only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
amr.harmony.mkrs_padj <- amr.harmony.mkrs[amr.harmony.mkrs$p_val_adj <0.01 ,]

list_feat<- c()
for(clust in unique(amr.harmony.mkrs_padj$cluster)){
  top_gene <- subset(amr.harmony.mkrs_padj,cluster==clust)$gene
  list_feat <- c(list_feat,head(top_gene[!top_gene %in% list_feat],3))
}

DotPlot(amr.harmony, features=list_feat, cols='RdBu') + RotatedAxis() + ggtitle("Marker genes") + ylab('Clusters') +theme(axis.text.x = element_text(angle = 90, hjust=1))

list_feat=c('MAL','GZMK', 'CCL5','DUSP2','HLA-DRA','LGALS1','CXCR3','HLA-DRB1','NKG7','GZMH','CST7','HIST1H1E','HIST1H1D','HIST1H1C','GNLY','GZMB','FGFBP2','FOXP3','IKZF2','RTKN2','NEAT1','XIST','MACF1','AL627171.2','PLCG2')

fig <- DotPlot(amr.harmony, features=list_feat, cols='RdBu') + RotatedAxis() + ggtitle("Marker genes") + ylab('Clusters') +theme(axis.text.x = element_text(angle = 90, hjust=1))

dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

ggsave(here::here("output", DOCNAME, "figure3-panelC.pdf"), fig,
        width = 9, height = 6, scale = 1)
ggsave(here::here("output", DOCNAME, "figure3-panelC.png"), fig,
        width = 9, height = 6, scale = 1)

fig
```

Panel E
------

```{r fig3-pE, fig.width = 12, fig.height = 10}
DefaultAssay(amr.harmony) <- 'SCT'
#Genes module autoreactivity
my_data <- read_excel(paste0(here::here("data/processed/amR053_List_features_genes_Update_11092024.xlsx")), sheet = "autoreact", col_names = F)
colnames(my_data) <- 'autoreact'
amr.harmony <- AddModuleScore(amr.harmony, features=list(my_data$autoreact), name='auto_reactivity', search=TRUE)
#Genes module H1N1
my_data <- read_excel(paste0(here::here("data/processed/amR053_List_features_genes_Update_11092024.xlsx")), sheet = "h1n1", col_names = F)
colnames(my_data) <- 'h1n1'
amr.harmony <- AddModuleScore(amr.harmony, features=list(my_data$h1n1), name='H1N1', search=TRUE)

# Feature plot auto reactivity highlight top 5%
top95p_auto <- tail(quantile(t(amr.harmony[[paste('auto_reactivity','1',sep='')]]),seq(0,1,0.05)),2)[1]
top_cells <- colnames(amr.harmony)[t(amr.harmony[[paste('auto_reactivity','1',sep='')]]) > top95p_auto]
pt_size_set <- rep(0.01, dim(amr.harmony)[2])
names(pt_size_set) <- colnames(amr.harmony)
pt_size_set[top_cells] <- 1
ft_modlist_auto <-FeaturePlot(amr.harmony, features=paste('auto_reactivity','1',sep=''), label=TRUE, pt.size=pt_size_set) + scale_colour_gradientn(colours=rev(brewer.pal(n=11, name="RdBu"))) + labs(title = "", subtitle = "Gene set score:\nauto-reactivity") + theme(plot.subtitle = element_text(size = 12))
#ft_modlist_auto

# Feature plot H1N1 highlight top 5%
top95p_h1n1 <- tail(quantile(t(amr.harmony[[paste('H1N1','1',sep='')]]),seq(0,1,0.05)),2)[1]
top_cells <- colnames(amr.harmony)[t(amr.harmony[[paste('H1N1','1',sep='')]]) > top95p_h1n1]
pt_size_set <- rep(0.01, dim(amr.harmony)[2])
names(pt_size_set) <- colnames(amr.harmony)
pt_size_set[top_cells] <- 1
ft_modlist_h1n1 <- FeaturePlot(amr.harmony, features=paste('H1N1','1',sep=''), label=TRUE, pt.size=pt_size_set) + scale_colour_gradientn(colours=rev(brewer.pal(n=11, name="RdBu"))) + labs(title = "", subtitle = "Gene set score:\nH1N1") +
  theme(plot.subtitle = element_text(size = 12))
#ft_modlist_h1n1

# List module score  violin Plot pairwise wilcoxon test
#AutoReactivity
liste_auto <- amr.harmony[[c('auto_reactivity1','seurat_clusters')]]
names(liste_auto) <- c('mod_score','clusters')
pwWlcx_auto <- pairwise_wilcox_test(liste_auto,formula=mod_score ~ clusters,p.adjust.method = "BH")
pwWlcx_auto_f =  pwWlcx_auto[pwWlcx_auto$group1 %in% c(2,6)| pwWlcx_auto$group2 %in% c(2,6), ]

mod_plot_auto <-  ggviolin(liste_auto, x='clusters', y='mod_score',palette=hue_pal()(10), fill='clusters') + 
stat_pvalue_manual(pwWlcx_auto_f, hide.ns = FALSE, y.position=0.5, step.increase=0.05) + ggtitle('Auto Reactivity')+ theme(legend.position = 'none') + xlab("scRNA-seq clusters") + ylab("Gene set score")

#H1N1
liste_h1n1 <- amr.harmony[[c('H1N11','seurat_clusters')]]
names(liste_h1n1) <- c('mod_score','clusters')
pwWlcx_h1n1 <- pairwise_wilcox_test(liste_h1n1,formula=mod_score ~ clusters,p.adjust.method = "BH")
pwWlcx_h1n1_f =  pwWlcx_h1n1[pwWlcx_h1n1$group1 %in% c(3,5)| pwWlcx_h1n1$group2 %in% c(3,5), ]


mod_plot_h1n1 <-  ggviolin(liste_h1n1, x='clusters', y='mod_score',palette=hue_pal()(10), fill='clusters') + 
stat_pvalue_manual(pwWlcx_h1n1_f, hide.ns = FALSE, y.position=0.5, step.increase=0.05) + ggtitle('H1N1') + theme(legend.position = 'none') + xlab("scRNA-seq clusters") + ylab("Gene set score")


fig <- plot_grid(ft_modlist_auto,ft_modlist_h1n1, mod_plot_auto, mod_plot_h1n1,  nrow = 2, labels = "E")

dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

ggsave(here::here("output", DOCNAME, "figure3-panelE.pdf"), fig,
        width = 12, height = 10, scale = 1)
ggsave(here::here("output", DOCNAME, "figure3-panelE.png"), fig,
        width = 12, height = 10, scale = 1)

fig
```

Supplementary figure 10
=======

```{r suppfig10, fig.width = 9, fig.height = 10}

#Indentify marker genes for each cluster
amr.harmony.mkrs_padj %>% group_by(cluster) %>% top_n(n = 5, wt = avg_log2FC) -> top5

#amr.harmony.mkrs_padj %>% group_by(cluster) %>% top_n(n = 5, wt = avg_log2FC) -> top5
#heatmap.mkrs <- DoHeatmap(amr.harmony, features=top5$gene, group.by='seurat_clusters')



# fig <- DoHeatmap(amr.harmony, features=top5$gene, group.by='seurat_clusters', assay = 'SCT', label = F)
# more than 30k cells, https://github.com/satijalab/seurat/issues/2724
DoHeatmap(subset(amr.harmony, downsample = 5000), 
          features=top5$gene,  group.by='seurat_clusters',
                  assay = "SCT")

fig <- DoHeatmap(subset(amr.harmony, downsample = 5000), 
          features=list_feat,  group.by='seurat_clusters',
                  assay = "SCT")


dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

ggsave(here::here("output", DOCNAME, "suppfigure10.pdf"), fig,
        width = 9, height = 10, scale = 1)
ggsave(here::here("output", DOCNAME, "suppfigure10.png"), fig,
        width = 9, height = 10, scale = 1)

fig
```

Supplementary figure 11
=======

```{r suppfig11, fig.width = 10, fig.height = 9}
Idents(amr.harmony) <- 'predicted.celltype.l2'
fig <- DimPlot(amr.harmony, split.by='predicted.celltype.l2',ncol=4,seed=42)

dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

ggsave(here::here("output", DOCNAME, "suppfigure11.pdf"), fig,
        width = 10, height = 9, scale = 1)
ggsave(here::here("output", DOCNAME, "suppfigure11.png"), fig,
        width = 10, height = 9, scale = 1)

fig
```


Supplementary figure 12
=======

```{r suppfig12, fig.width = 12, fig.height = 10}
ft_modlist = list()
mod_plot_list = list()
for (sname in c("Poon_2022_Lung_CD4_TRM","Poon_2022_Skin_CD4_TRM","Poon_2022_jejunum_CD4_TRM", 
                "Zhang_2022_Liver_HBV_CXCL13_CD4", "Zheng_2017_Liver_HCC_CXCL13_CD4", "Argyriou_2022_Synovial_TPH", "Law_2024_Synovial_TPH")){
  
#Genes module 
my_data <- read_excel(paste0(here::here("data/processed/amR053_List_features_genes_Update_11092024.xlsx")), sheet = sname, col_names = F)
colnames(my_data) <- sname
amr.harmony <- AddModuleScore(amr.harmony, features=list(my_data[[sname]]), name= sname, search=TRUE)

# Feature plot highlight top 5%
top95p <- tail(quantile(t(amr.harmony[[paste(sname,'1',sep='')]]),seq(0,1,0.05)),2)[1]
top_cells <- colnames(amr.harmony)[t(amr.harmony[[paste(sname,'1',sep='')]]) > top95p]
pt_size_set <- rep(0.01, dim(amr.harmony)[2])
names(pt_size_set) <- colnames(amr.harmony)
pt_size_set[top_cells] <- 1
ft_modlist[[sname]] <-FeaturePlot(amr.harmony, features=paste(sname,'1',sep=''), pt.size=pt_size_set) + scale_colour_gradientn(colours=rev(brewer.pal(n=11, name="RdBu"))) + labs(title = "", subtitle = paste0("Gene set score:\n",sname)) + theme(plot.subtitle = element_text(size = 12)) & NoLegend() & NoAxes()


# List module score  violin Plot pairwise wilcoxon test
liste_tempo <- amr.harmony[[c(paste(sname,'1',sep=''),'seurat_clusters')]]
names(liste_tempo) <- c('mod_score','clusters')
pwWlcx_auto <- pairwise_wilcox_test(liste_tempo,formula=mod_score ~ clusters,p.adjust.method = "BH")
pwWlcx_auto_f =  pwWlcx_auto[pwWlcx_auto$group1 %in% c(2,6)| pwWlcx_auto$group2 %in% c(2,6), ]

mod_plot_list[[sname]] <-  ggviolin(liste_tempo, x='clusters', y='mod_score',palette=hue_pal()(10), fill='clusters') + 
stat_pvalue_manual(pwWlcx_auto_f, hide.ns = FALSE, y.position=0.5, step.increase=0.05) + ggtitle(sname)+ theme(legend.position = 'none') + xlab("scRNA-seq clusters") + ylab("Gene set score")

}

fig <- plot_grid(mod_plot_list[[1]], ft_modlist[[1]],
                 mod_plot_list[[2]], ft_modlist[[2]],
                 mod_plot_list[[3]], ft_modlist[[3]],
                 mod_plot_list[[4]], ft_modlist[[4]],
                 mod_plot_list[[5]], ft_modlist[[5]],
                 mod_plot_list[[6]], ft_modlist[[6]],
                 mod_plot_list[[7]], ft_modlist[[7]],
                 nrow = 3, labels = "A")

dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

ggsave(here::here("output", DOCNAME, "suppfigure12.pdf"), fig,
        width = 18, height = 12, scale = 1)
ggsave(here::here("output", DOCNAME, "suppfigure12.png"), fig,
        width = 18, height = 12, scale = 1)

fig
```


Output files
==============

This table describes the output files produced by this document. Right click
and _Save Link As..._ to download the results.

```{r save}
saveRDS(amr.harmony, here::here("data/processed/05-sc_seurat.rds"))

```


```{r output}
dir.create(here::here("output", DOCNAME), showWarnings = FALSE)

knitr::kable(data.frame(
     File = c(
         getDownloadLink2("figure3-panelB.png", str_sub(here::here("output", DOCNAME),2,-1)),
         getDownloadLink2("figure3-panelB.pdf", str_sub(here::here("output", DOCNAME),2,-1)),
         getDownloadLink2("figure3-panelC.png", str_sub(here::here("output", DOCNAME),2,-1)),
         getDownloadLink2("figure3-panelC.pdf", str_sub(here::here("output", DOCNAME),2,-1)),
         getDownloadLink2("figure3-panelE.png", str_sub(here::here("output", DOCNAME),2,-1)),
         getDownloadLink2("figure3-panelE.pdf", str_sub(here::here("output", DOCNAME),2,-1)),
         getDownloadLink2("suppfigure10.png", str_sub(here::here("output", DOCNAME),2,-1)),
         getDownloadLink2("suppfigure10.pdf", str_sub(here::here("output", DOCNAME),2,-1)),
         getDownloadLink2("suppfigure11.png", str_sub(here::here("output", DOCNAME),2,-1)), 
         getDownloadLink2("suppfigure11.pdf", str_sub(here::here("output", DOCNAME),2,-1)),
         getDownloadLink2("suppfigure12.png", str_sub(here::here("output", DOCNAME),2,-1)), # bug one "/" plus
         getDownloadLink2("suppfigure12.pdf", str_sub(here::here("output", DOCNAME),2,-1))
     ),
     Description = c(
         "Figure 3 panel B (PNG)",
         "Figure 3 panel B (PDF)",
         "Figure 3 panel C (PNG)",
         "Figure 3 panel C (PDF)",
         "Figure 3 panel E (PNG)",
         "Figure 3 panel E (PDF)",
         "Supplementary Figure 10 (PNG)",
         "Supplementary Figure 10 (PDF)",
         "Supplementary Figure 11 (PNG)",
         "Supplementary Figure 11 (PDF)",
         "Supplementary Figure 12 (PNG)",
         "Supplementary Figure 12 (PDF)"
     )
 ))
```


Session information
==============

```{r session-info, cache = FALSE}
devtools::session_info()
```
